Technique for \earlywarning\ detection of compact binary coalescence and its implication for multimessenger astronomy

### 2011-06-02 results.tex
some more text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@828 ccdd05d5-56c1-461a-bf98-85f6876523b4

We expect two contributions to the \SNR\ loss to arise in our implementation of
the \lloid\ algorithm.  The first is the \SNR\ loss due to the truncation of
the SVD basis, which is fundamental, and estimates for it
exist~\cite{Cannon:2010p10398}.  The second comes from non ideal
implementations of resampling that cause signal loss.  We have measured the
effect of the resampling in the pipeline alone on a single waveform to gauge
the magnitude of the loss.  It is shown in figure \ref{fig:resamp_mm} as a
function of the quality of the {\tt audioresample} element (which is
proportional to the filter length used).  In both cases we have to compare the
\SNR\ loss to the expected \SNR\ loss that arises from the discreteness of the
template bank which is typically $\sim 3\%$.  We will consider a comparable
\SNR\ loss to be tolerable, but ideally would like to have a factor of 10 lower
loss from the \lloid\ implementation (i.e. no more than $\sim 0.3 \%$).  
%
\begin{figure}
\includegraphics{resamp_mm.pdf}
\caption{\label{fig:resamp_mm}Mis-match as a function of the {\tt
audioresample} element's quality parameter.  Increasing the quality changes the
length of the sinc function filter that is used to decimate and interpolate the
data and filter output in our test pipeline.  Choosing a quality of 4
corresponds to a filter length of 64 samples and results in an acceptable loss
of .2 \% in \SNR\ for the chosen waveform.}
\end{figure}

\begin{figure}
	\label{fig:hist}
	\begin{center}
		\includegraphics{bw.pdf}
		\caption{Box-and-whisker plot of mismatch between nominal template bank and \lloid\ measured impulse respones.  The \textsc{svd} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.  The whiskers denote the minimum and maximum mismatch over all templates.}
	\end{center}
\end{figure}

@@

@@

%\FIXME{
%I would like to see this section changed to emphasize the low-latency time
%domain filtering 1. Use a live simulated white noise source (this ignores the
%latency of whitening, but that goes beyond the scope of this paper, and we
%mention this and perhaps suggest some exporation 2. Use TD filtering of N
%templates 3. Present the performance and latency and provide estimates for
%number of cores required for realtime ALIGO search based on current
%infrastructure 4. Compute the impulse responses of the templates and histogram
%the SNR loss for the SVD and time/slice/resampling 5. Put  a tee after some of
%the low frequency stages and perhaps test the possibility of predictive
%filtering (also with latency measurments) This will remove the complications of
%running a full pipeline.  We need to sort things out before we do that, and
%this paper should not be delayed any further.  The above tests will make the
%point we need to make
%}

%\subsection{Detector noise characteristics}

%We tested the new detection method with mock Advanced \textsc{ligo} data having a power spectrum prescribed by the ``zero detuning, high power'' noise model in \cite{Shoemaker:2009p9770}.

>>>>

We expect two contributions to the \SNR\ loss to arise in our implementation of
the \lloid\ algorithm.  The first is the \SNR\ loss due to the truncation of
the SVD basis, which is fundamental, and estimates for it
exist~\cite{Cannon:2010p10398}.  The second comes from non ideal
implementations of resampling that cause signal loss.  We have measured the
effect of the resampling in the pipeline alone on a single waveform to gauge
the magnitude of the loss.  It is shown in figure \ref{fig:resamp_mm} as a
function of the quality of the {\tt audioresample} element (which is
proportional to the filter length used).  In both cases we have to compare the
\SNR\ loss to the expected \SNR\ loss that arises from the discreteness of the
template bank which is typically $\sim 3\%$.  We will consider a comparable
\SNR\ loss to be tolerable, but ideally would like to have a factor of 10 lower
loss from the \lloid\ implementation (i.e. no more than $\sim 0.3 \%$).  
%
\begin{figure}
\includegraphics{resamp_mm.pdf}
\caption{\label{fig:resamp_mm}Mis-match as a function of the {\tt
audioresample} element's filter length.  Increasing the filter length changes
the length of the sinc function filter that is used to decimate and interpolate
the data and filtered time series in our test pipeline.  Choosing a filter
length of 64 samples results in an acceptable loss of .2 \% in \SNR\ for the
chosen waveform.}
\end{figure}

\begin{figure}
	\label{fig:hist}
	\begin{center}
		\includegraphics{bw.pdf}
		\caption{Box-and-whisker plot of mismatch between nominal
template bank and \lloid\ measured impulse respones.  The \textsc{svd}
tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$.
The whiskers denote the minimum and maximum mismatch over all templates.  The
diminishing improvement is caused by hitting the resampling match limit.}
	\end{center}
\end{figure}

@@

In order to check the latency of our implementation of the \lloid\ algorithm we
used the pipeline described in \ref{fig:pipeline} with a live source connected
to the input.  We then measured the timestamp of the live source and the output
of the final \SNR\ time series.  Since \gstreamer\ is not a
sample-in-sample-out realtime infrastructure, it does work by passing small
buffers of data and filtered data through the pipeline.  We set the buffer size
on the input to be 1/4 s.  In an ideal case the latency should be nearly 1/4 s.
However, it is necessary to queue things in the pipeline to assure that there
are no issues with threading. Sometimes buffers get delayed.  We measured the
average filtering latency to be \FIXME{1 s}.

@@

Using the pipeline described previously, we also measured the computational
cost.  We used an 8 socket, quad core AMD\texttrademark\ machine operating at
2.7 GHz per core.  We found that filtering \FIXME{657} templates in real time
required about \FIXME{6} of the available cores.  We thus conclude that we
require about 1 core for every \FIXME{100} templates.  We note that this is
well within the computational limits of a modern, computing cluster assuming
that the scaling holds for the entire mass range of interest, we would need
about \FIXME{300} cores per detector.  This requirment should be even less of a
burden in the advanced detecor era, when presumably clusters built in that era
will have a factor of 4 or 8 more cores than the ones today.
